```{r}
#Setting directory
setwd("C:/Users/Ique Nepomuceno/Documents/Performance Marketing/SARIMA Model")
```
```{r}
#Loading Packages
library("readxl")
library("tseries")
library("forecast")
library("openxlsx")
library("ggplot2")
```

```{r}
# Loading Data Frames
# All the needed data, including the R notebook should be in folder for easier coding/operation. 
# Monthly_rev contains daily raw data of brands. 

#Reading file
monthly_rev <- read_excel("C:\\Users\\Ique Nepomuceno\\Documents\\Performance Marketing\\SARIMA Model\\BH DEFR Data_v2.xlsx", 
               sheet=1, 
               col_names=TRUE, 
               col_types=NULL,
               skip=0)

attach(monthly_rev)
summary(Revenue)
```
```{r}
# Preparing for the Model

#Splitting Data into Test & Train Data
#Splitting 2015-2018 vs 2019
#sd = Start Date or First Fiscal Year data available

train<-head(monthly_rev, nrow(monthly_rev)-12)
test<-tail(monthly_rev, 12)
sd=2018
```

```{r}

#converting data into time series format
#converting split data into time series format
#The '$' symbol lets you select particular columns in a table or dataframe

#Revenue
tsData= ts(monthly_rev$Revenue, start = c(sd,2), frequency = 12)
tsTrain=ts(train$Revenue, start=c(sd,2), frequency =12)
tsTest=ts(test$Revenue, start=c(2019,2), frequency =12)
autoplot(tsTrain) + autolayer(tsTest)
autoplot(tsData)

#Conversions
tsDataCon = ts(monthly_rev$Conversions, start = c(sd,2), frequency = 12)
tsTrainCon = ts(train$Conversions, start=c(sd,2), frequency =12)
tsTestCon = ts(test$Conversions, start=c(2019,2), frequency =12)

#Cost
tsDataCost = ts(monthly_rev$Cost, start = c(sd,2), frequency = 12)
tsTrainCost = ts(train$Cost, start=c(sd,2), frequency =12)
tsTestCost = ts(test$Cost, start=c(2019,2), frequency =12)

#Clicks
tsDataClicks = ts(monthly_rev$Clicks, start = c(sd,2), frequency = 12)
tsTrainClicks = ts(train$Clicks, start=c(sd,2), frequency =12)
tsTestClicks = ts(test$Clicks, start=c(2019,2), frequency =12)

#Impressions
tsDataImp = ts(monthly_rev$Impressions, start = c(sd,2), frequency = 12)
tsTrainImp = ts(train$Impressions, start=c(sd,2), frequency =12)
tsTestImp = ts(test$Impressions, start=c(2019,2), frequency =12)

```
## Finding the Appropriate pdqPDQm Values 

**Trend Elements**       
p: Trend autoregression order    
d: Trend difference order   
q: Trend moving average order    

**Seasonal Elements**      
P: Seasonal autoregressive order     
D: Seasonal difference order        
Q: Seasonal moving average order      
m: The number of time steps for a single seasonal period     

Source: https://machinelearningmastery.com/sarima-for-time-series-forecasting-in-python/ 

```{r}
#Finding the appropriate pdqPDQ values of each dataset

arima<-auto.arima(tsData,seasonal=TRUE)
arimaCon<-auto.arima(tsDataCon,seasonal=TRUE)
arimaCost<-auto.arima(tsDataCost,seasonal=TRUE)
arimaClicks<-auto.arima(tsDataClicks,seasonal=TRUE)
arimaImp<-auto.arima(tsDataImp,seasonal=TRUE)

#Finding the appropriate pdqPDQ values for each training dataset
arimaTrain<-auto.arima(tsTrain,seasonal=TRUE)
arimaTrainCon<-auto.arima(tsTrainCon, seasonal=TRUE)
arimaTrainCost<-auto.arima(tsTrainCost, seasonal=TRUE)
arimaTrainClicks<-auto.arima(tsTrainClicks, seasonal=TRUE)
arimaTrainImp<-auto.arima(tsTrainImp, seasonal=TRUE)
```

## Using setNames for easier identification
Remember this Pattern: p q P Q m d D
```{r}
#using setNames for easier identification:
#remember this pattern: p q P Q m d D
# arimaTrain$arma

input<-setNames(arima$arma, c("p","q","P","Q","m","d","D"))
inputCon<-setNames(arimaCon$arma, c("p","q","P","Q","m","d","D"))
inputCost<-setNames(arimaCost$arma, c("p","q","P","Q","m","d","D"))
inputClicks<-setNames(arimaClicks$arma, c("p","q","P","Q","m","d","D"))
inputImp<-setNames(arimaImp$arma, c("p","q","P","Q","m","d","D"))

inputTrain<-setNames(arimaTrain$arma, c("p","q","P","Q","m","d","D"))
inputTrainCon<-setNames(arimaTrainCon$arma, c("p","q","P","Q","m","d","D"))
inputTrainCost<-setNames(arimaTrainCost$arma, c("p","q","P","Q","m","d","D"))
inputTrainClicks<-setNames(arimaTrainClicks$arma, c("p","q","P","Q","m","d","D"))
inputTrainImp<-setNames(arimaTrainImp$arma, c("p","q","P","Q","m","d","D"))

```

# Fitting the Model
```{r}
#fitting the model
fitARIMA <- arima(tsData, 
                  order=c(input["p"],input["d"],input["q"]),
                  seasonal = list(order = c(input["P"],input["D"],input["Q"]),
                  period = 12),
                  method="ML")

fitARIMACon <- arima(tsDataCon,
                     order=c(inputCon["p"],inputCon["d"],inputCon["q"]), 
                     seasonal = list(order = 
                     c(inputCon["P"],inputCon["D"],inputCon["Q"]),
                     period = 12),
                     method="ML")

fitARIMACost <- arima(tsDataCost,  
                    order=c(inputCost["p"],inputCost["d"],inputCost["q"]),
                    seasonal = list(order = 
                    c(inputCost["P"],inputCost["D"],inputCost["Q"]), 
                    period = 12), 
                    method="ML")

fitARIMAClicks <- arima(tsDataClicks, 
                  order=c(inputClicks["p"],inputClicks["d"],inputClicks["q"]), 
                  seasonal = list(order = 
                  c(inputClicks["P"],inputClicks["D"],inputClicks["Q"]), 
                  period = 12), 
                  method="ML")

fitARIMAImp <- arima(tsDataImp, 
                     order=c(inputImp["p"],inputImp["d"],inputImp["q"]), 
                     seasonal = list(order = 
                     c(inputImp["P"],inputImp["D"],inputImp["Q"]), 
                     period = 12), 
                     method="ML")

#fitting the training models
#Revenue
fitTrain <- arima(tsTrain,
                  order=c(inputTrain["p"],inputTrain["d"],inputTrain["q"]),
                  seasonal = list(order =
                  c(inputTrain["P"],inputTrain["D"],inputTrain["Q"]),
                  period = 12),
                  method="ML")
 
# #Conversions
fitTrainCon <- arima(tsTrainCon,
             order=c(inputTrainCon["p"],inputTrainCon["d"],inputTrainCon["q"]),
             seasonal = list(order =
             c(inputTrainCon["P"],inputTrainCon["D"],inputTrainCon["Q"]),
             period = 12),
             method="ML")

# #Cost
fitTrainCost <- arima(tsTrainCost,
          order=c(inputTrainCost["p"],inputTrainCost["d"],inputTrainCost["q"]),
          seasonal = list(order =
          c(inputTrainCost["P"],inputTrainCost["D"],inputTrainCost["Q"]),
          period = 12),
          method="ML")

# #Clicks
fitTrainClicks <- arima(tsTrainClicks,
    order=c(inputTrainClicks["p"],inputTrainClicks["d"],inputTrainClicks["q"]),
    seasonal = list(order =
    c(inputTrainClicks["P"],inputTrainClicks["D"],inputTrainClicks["Q"]),
    period = 12),
    method="ML")

# #Impressions
fitTrainImp <- Arima(tsTrainImp,
            order=c(inputTrainImp["p"],inputTrainImp["d"],inputTrainImp["q"]),
            seasonal = list(order =
            c(inputTrainImp["P"],inputTrainImp["D"],inputTrainImp["Q"]),
            period = 12),
            method="ML")

```

```{r}
#testing the model
# acf(residuals(fitARIMA))
# pacf(residuals(fitARIMA))
# Box.test(residuals(fitARIMA)
         
#testing the training model
# acf(residuals(fitARIMATrain))
# pacf(residuals(fitARIMATrain))
# Box.test(residuals(fitARIMATrain))
```

# Forecasting Values

## Forecasts
There are other tests that can be conducted after this. 
```{r}
#forecasting the value -- there are tests that needs to be done again after this
futureVal<-forecast(fitARIMA,h=12,level=c(99.5))

futureValCon <-forecast(fitARIMACon,h=12,level=c(99.5))
futureValCost <-forecast(fitARIMACost,h=12,level=c(99.5))
futureValClicks <-forecast(fitARIMAClicks,h=12,level=c(99.5))
futureValImp <-forecast(fitARIMAImp,h=12,level=c(99.5))

#forecasting value from training model
futureTrain<-forecast(fitTrain,h=12,level=c(99.5))

futureTrainCon <-forecast(fitTrainCon,h=12,level=c(99.5))
futureTrainCost <-forecast(fitTrainCost,h=12,level=c(99.5))
futureTrainClicks <-forecast(fitTrainClicks,h=12,level=c(99.5))
futureTrainImp <-forecast(fitTrainImp,h=12,level=c(99.5))
```

```{r}
#checking the result we have in graph format:
#To see the values
plot(futureVal)
plot(futureValClicks)
plot(futureValCost)
plot(futureValCon)
plot(futureValImp)

#To see training model's forecasts and the actual values
autoplot(futureTrain) + autolayer(tsTest)
```

## Testing Accuracy of Forecast Model and Test Model
```{r}
#Accuracy Metrics of Forecast Model
Accu=accuracy(futureVal)
AccuCon=accuracy(futureValCon)
AccuCost=accuracy(futureValCost)
AccuClicks=accuracy(futureValClicks)
AccuImp=accuracy(futureValImp)

#Accuracy Metrics of Test Models
AccuTrain = accuracy(futureTrain, tsTest)
AccuTrainCon=accuracy(futureTrainCon, tsTestCon)
AccuTrainCost=accuracy(futureTrainCost, tsTestCost)
AccuTrainClicks=accuracy(futureTrainClicks, tsTestClicks)
AccuTrainImp=accuracy(futureTrainImp, tsTestImp)
print(futureVal)
print(futureValCon)
print(futureValCost)
print(futureValClicks)
print(futureValImp)
print(futureTrain)
```

```{r}
#To write the data into a xlsx file
#xlsx file must already exist in directory. filename here is test.xlsx
#write.xlsx command comes from "openxlsx" package
#You can choose to change the name of the file test to something else

#To write the data into a csv file
#csv file must already exist in directory. filename here is test.csv
#write.csv(futureVal, file="test.csv")
#You can choose to change the name of the file test to something else

#Building a workbook to contain both forecasts and actual values
testData<-createWorkbook()
addWorksheet(testData, "Revenue")
addWorksheet(testData, "Conversions")
addWorksheet(testData, "Cost")
addWorksheet(testData, "Clicks")
addWorksheet(testData, "Impressions")

writeData(testData, "Revenue", futureVal, startCol = 7)
writeData(testData, "Revenue", Accu[4], startCol = 7, startRow = 16)
writeData(testData, "Revenue", Accu[5], startCol = 8, startRow = 16)
writeData(testData, "Revenue", "MPE", startCol = 7, startRow = 15)
writeData(testData, "Revenue", "MAPE", startCol = 8, startRow = 15)

writeData(testData, "Revenue", tsTest, startCol = 4)
writeData(testData, "Revenue", futureTrain)
writeData(testData, "Revenue", "MPE", startCol = 1, startRow = 15)
writeData(testData, "Revenue", "MAPE", startCol = 2, startRow = 15)
writeData(testData, "Revenue", AccuTrain[8], startCol = 1, startRow = 16)
writeData(testData, "Revenue", AccuTrain[10], startCol = 2, startRow = 16)

writeData(testData, "Conversions", futureValCon, startCol = 7)
writeData(testData, "Conversions", AccuCon[4], startCol = 7, startRow = 16)
writeData(testData, "Conversions", AccuCon[5], startCol = 8, startRow = 16)
writeData(testData, "Conversions", "MPE", startCol = 7, startRow = 15)
writeData(testData, "Conversions", "MAPE", startCol = 8, startRow = 15)

writeData(testData, "Conversions", tsTestCon, startCol = 4)
writeData(testData, "Conversions", futureTrainCon)
writeData(testData, "Conversions", "MPE", startCol = 1, startRow = 15)
writeData(testData, "Conversions", "MAPE", startCol = 2, startRow = 15)
writeData(testData, "Conversions", AccuTrainCon[8], startCol = 1, startRow = 16)
writeData(testData, "Conversions", AccuTrainCon[10], startCol = 2, startRow = 16)

writeData(testData, "Cost", futureValCost, startCol = 7)
writeData(testData, "Cost", AccuCost[4], startCol = 7, startRow = 16)
writeData(testData, "Cost", AccuCost[5], startCol = 8, startRow = 16)
writeData(testData, "Cost", "MPE", startCol = 7, startRow = 15)
writeData(testData, "Cost", "MAPE", startCol = 8, startRow = 15)

writeData(testData, "Cost", tsTestCost, startCol = 4)
writeData(testData, "Cost", futureTrainCost)
writeData(testData, "Cost", "MPE", startCol = 1, startRow = 15)
writeData(testData, "Cost", "MAPE", startCol = 2, startRow = 15)
writeData(testData, "Cost", AccuTrainCost[8], startCol = 1, startRow = 16)
writeData(testData, "Cost", AccuTrainCost[10], startCol = 2, startRow = 16)

writeData(testData, "Clicks", futureValClicks, startCol = 7)
writeData(testData, "Clicks", AccuClicks[4], startCol = 7, startRow = 16)
writeData(testData, "Clicks", AccuClicks[5], startCol = 8, startRow = 16)
writeData(testData, "Clicks", "MPE", startCol = 7, startRow = 15)
writeData(testData, "Clicks", "MAPE", startCol = 8, startRow = 15)

writeData(testData, "Clicks", tsTestClicks, startCol = 4)
writeData(testData, "Clicks", futureTrainClicks)
writeData(testData, "Clicks", "MPE", startCol = 1, startRow = 15)
writeData(testData, "Clicks", "MAPE", startCol = 2, startRow = 15)
writeData(testData, "Clicks", AccuTrainClicks[8], startCol = 1, startRow = 16)
writeData(testData, "Clicks", AccuTrainClicks[10], startCol = 2, startRow = 16)

writeData(testData, "Impressions", futureValImp, startCol = 7)
writeData(testData, "Impressions", AccuImp[4], startCol = 7, startRow = 16)
writeData(testData, "Impressions", AccuImp[5], startCol = 8, startRow = 16)
writeData(testData, "Impressions", "MPE", startCol = 7, startRow = 15)
writeData(testData, "Impressions", "MAPE", startCol = 8, startRow = 15)

writeData(testData, "Impressions", tsTestImp, startCol = 4)
writeData(testData, "Impressions", futureTrainImp)
writeData(testData, "Impressions", "MPE", startCol = 1, startRow = 15)
writeData(testData, "Impressions", "MAPE", startCol = 2, startRow = 15)
writeData(testData, "Impressions", AccuTrainImp[8], startCol = 1, startRow = 16)
writeData(testData, "Impressions", AccuTrainImp[10], startCol = 2, startRow = 16)


#Filename should be changed before running if you'd like to retain previous files
saveWorkbook(testData, "BHUK Forecasts_test3.xlsx")
```
